---
title: "Week-2"
author: "Rajesh Kalakoti"
date: "2023-08-03"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
  html_document: default
  word_document: default
header-includes:
  - \usepackage[linesnumbered,ruled,lined,boxed]{algorithm2e}
  - \usepackage{amsmath}
  - \usepackage{amsfonts}
  - \usepackage{placeins}
  - \usepackage{float} # For math fonts like \mathbb
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reticulate)
library(here)
```

* Packages
    + [devtools](https://www.r-project.org/nosvn/pandoc/devtools.html) 
    + [tidyverse](https://www.tidyverse.org/packages/)
    + here

```{r}
library(here)
project_path <- here()
source(here("R","utils.R"))
source(here("R","distance_functions.R"))
```

```{r}
data(iris)
iris_data <- iris[, 1:4]  # Selecting the first 4 columns (features)
# Calculate the Hopkins statistic using the hopkins_stat function
?hopkins_stat
hopkins_value <- hopkins_stat(iris_data)
print(hopkins_value)
```

# Clustering

Given a clustering \(C = \{C_1, C_2, \ldots, C_k\}\), we need some scoring function that evaluates its quality or goodness. This sum of squared errors scoring function is defined as:
\[ W(C) = \frac{1}{2} \sum_{k=1}^{K} \sum_{i: C(i)=k} \|x_i - \bar{x}_k\|^2 \]
The goal is to find the clustering that minimizes:
\[ C^* = \arg \min_C \{ W(c) \} \]
K-means employs a greedy iterative approach to find a clustering that minimizes loss function. 
\begin{algorithm}
\LinesNumbered % Add line numbers to the algorithm
\caption{K-means Algorithm}
\KwData{$D, k, \varepsilon$}

\textbf{K-means}($D, k, \varepsilon$): {
  
  $t \leftarrow 0$\;
  Randomly initialize $k$ centroids: $\mu_{1}^{t}, \mu_{2}^{t}, \ldots, \mu_{n}^{t} \in \mathbb{R}^d$\;
  \Repeat{$\sum_{i=1}^{k} ||\mu_{i}^t - \mu_{i}^{t-1}||^2 \le \varepsilon $}{
    $t \leftarrow t+1$\;
    $C_i \leftarrow \emptyset$ for all \(i = 1, \ldots, k\)
    
    /* Cluster assignment step */

    
    \For{$x_j \in D$}{
      $i^* \leftarrow \text{argmin}_i \{||x_j - \mu_{i}^{t-1}||^2 \}$\;
      /* assign $x_j$ to closest centroid */
      
      $C_{i^*} \leftarrow C_{i^*} \cup \{x_j\}$\;
    }
    
        
    \For{$i = 1, .., k $}{
      $  \mu_{i}^{t} \leftarrow \frac{1}{|C_i|} \sum_{x_j \in C_i} X_j $
    }
  }
}
\end{algorithm}



```{r}



```

