---
title: "week1"
author: "Rajesh Kalakoti"
date: "2023-08-03"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
  html_document: default
  word_document: default
header-includes:
  - \usepackage[linesnumbered,ruled,lined,boxed]{algorithm2e}
  - \usepackage{amsmath}  # For math symbols like \mathbb
  - \usepackage{amsfonts} # For math fonts like \mathbb
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reticulate)
```

* Packages
    + [devtools](https://www.r-project.org/nosvn/pandoc/devtools.html) 
    + [tidyverse](https://www.tidyverse.org/packages/) 
        - sub-sub-item 1



## Clustering

Given a clustering \(C = \{C_1, C_2, \ldots, C_k\}\), we need some scoring function that evaluates its quality or goodness. This sum of squared errors scoring function is defined as:
\[ W(C) = \frac{1}{2} \sum_{k=1}^{K} \sum_{i: C(i)=k} \|x_i - \bar{x}_k\|^2 \]

The goal is to find the clustering that minimizes:

\[ C^* = \arg \min_C \{ W(c) \} \]

K-means employs a greedy iterative approach to find a clustering that minimizes loss function. 


## Algorithm 13.1: K-means Algorithm
**K-means** (*D*, *k*, *ε*):
1. Initialize \(t = 0\). Randomly initialize \(k\) centroids: \(\mu_{t1},
\mu_{t2}, \ldots, \mu_{tk} \in \mathbb{R}^d\). 3. **repeat** 4.   \(t \leftarrow
t + 1\).
   **// Cluster Assignment Step**
5.   **foreach** \(x_j \in D\) **do** 6.     \(j^* \leftarrow \arg \min_i \|x_j
- \mu_{ti}\|^2\).
   **// Assign \(x_j\) to closest centroid**
7.     \(C_{j^*} \leftarrow C_{j^*} \cup \{x_j\}\).
   **// Centroid Update Step**
8.   **foreach** \(i = 1\) **to** \(k\) **do** 9.     \(\mu_{ti} \leftarrow
\frac{1}{|C_i|} \sum_{x_j \in C_i} x_j\).
     **until** \(\sum_{i=1}^k \|\mu_{ti} - \mu_{t-1i}\|^2 \leq \varepsilon\).



\begin{algorithm}
\LinesNumbered % Add line numbers to the algorithm
\caption{K-means Algorithm}
\KwData{$D, k, \varepsilon$}
\KwResult{Result $y$}

\textbf{K-means}($D, k, \varepsilon$) {
  
  $t \leftarrow 0$\;
  Randomly initialize $k$ centroids: $\mu_{1}^{t}, \mu_{2}^{t}, \ldots, \mu_{n}^{t} \in \mathbb{R}^d$\;
  \Repeat{termination condition}{
    $t \leftarrow t+1$\;
    /* Cluster assignment step */
    
    \For{$x_j \in D$}{
      $j^* \leftarrow \text{argmin}_i \{||x_j - \mu_{i}^t||^2 \}$\;
      /* assign $x_j$ to closest centroid */
      
      $C_{j^*} \leftarrow C_{j^*} \cup \{x_j\}$\;
    }
  }
  Perform additional steps\;
  \For{$i = 1$ to $n$}{
    Update $y \leftarrow y \times x$\;
  }
  \textbf{return} $y$\;
}
\end{algorithm}




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
?entropy
```


```{r}

```


